---
title: "assignment3"
author: "Vanessa Löw"
date: "2023-03-15"
output: html_document
---

```{r}
suppressPackageStartupMessages({
  library(AnnotationHub)
  library(Rsubread)
  library(rtracklayer)
  library(Biostrings)
  library(Rfastp)
  library(epiwraps)
})
ah <- AnnotationHub()
```

```{r}
getwd()
```

*1. Download the following Drosophila ChIP-seq for the protein CTCF:

```{r, eval=FALSE}
#increase download times for large downloads
options(timeout=3600)
#creation of subdirectory
dir.create("raw")
download.file("https://www.encodeproject.org/files/ENCFF127RRR/@@download/ENCFF127RRR.fastq.gz", dest="raw/Myc.fastq.gz")
```
*2. Process it from the raw data, obtaining:
**a. bam file

#Aligning reads to create .bam files

## Using Rsubread

### Building a genome index for mapping

This we have to do only once for a genome, and can then re-use across projects
the trimfastq files look exactly like the normal fastq files, but are smaller

```{r, eval=FALSE}
# we get the genome sequence from AnnotationHub
genome <- ah[["AH49674"]]
# we create a new directory that will contain the genome index
dir.create("BDGP6_genome")
# we write the genome sequence in fasta format
export(import.2bit(genome), "BDGP6_genome/genome.fasta.gz", compress=TRUE)
# we build a Rsubread index (can also be indexed by other software of choice)
Rsubread::buildindex("BDGP6_genome/rsubread", reference="BDGP6_genome/genome.fasta.gz")
```

### Trimming
especially important in epigenomics data, not so much in RNA data
## Using Rfastp
(no input control needed for this exercise)

```{r}
dir.create("aligned")
# package::function specifies that a certain function is from a certain package (for functions with identical names)
align.stats <- Rsubread::align(index="BDGP6_genome/rsubread", type="dna",
                               readfile1=c("rfastp.trimmed/Myc_R1.fastq.gz"),
                               output_file=c("aligned/Myc.bam"),
                               nthreads=6, sortReadsByCoordinates=TRUE)
#alignment statistics e.g., total reads, unmapped reads etc.
align.stats
```

**b. peaks
## Using MACS2
```{bash engine.opts='-l'}
mkdir -p peaks
macs2 callpeak --outdir peaks -n Myc --gsize dm -t aligned/Myc.bam -c aligned/input.bam
# macs3 works in the same way
```

Import the peaks in R:

```{r}
peaks <- rtracklayer::import("peaks/Myc_peaks.narrowPeak")
source("peaks/Myc_model.r")
peaks
```
*** I do not get any objects here, as Myx_peaks.narrowPeak is an empty file...

*3. Report:
**a. how many reads (and what percentage) were mapped
2351375 out of 3504769 reads were mapped, which is equivalent to 66%
**b. how many peaks were found
Number of peaks found is 1696 (paired)
*4. Plot the signal around one of the peaks
#I cannot plot the signal as the file is empty so there is nothing to plot, so I had to stop this function from running to be able to knit the document
```{r, eval=FALSE}
library(epiwraps)
plotSignalTracks(peaks)
```